Secure Composition of Robust and Optimising Compilers

### 2023-07-12 popl/main.tex
nits

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
% \MPpost{
%   this abstract fails in setting up the robustness argument.
%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
% }
% context
To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety as well as cryptographic constant time.
% need
Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
% task
Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property the composition of those secure compiler passes preserve.
%there is no way to tell what kind of security property will the composition of those secure compilers preserve.

% object
This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
% findings
Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a secure multi-pass compiler that preserves those security-relevant properties.
% conclusion
Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
% 
From an engineering perspective, this is the desirable approach to building secure compilers.

@@

\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
Together, \gls*{sms}, \gls*{tms} and \gls*{scct}, a variant of \gls*{cct}, yield \gls*{msscct}, which is the gold standard of security properties for secure applications: programs attaining \gls*{msscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels.
As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the various architectural details where their code will run.

@@

Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
Thus, robust compilers can be used to prevent, e.g., vulnerabilities resulting from uninitialised memory, e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
\end{example}

@@

@@

Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the concept of satisfaction can be extended with {\em robustness}.
Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has.
%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
This translation is considered {\em correct} if it is semantics-preserving.
That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ in such a way that they are ,,compatible''.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarises.

@@

\begin{example}[Types]
  Suppose $\src{L}$ is a statically typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
  Consider the following $\src{L}$ component and its compiled version below.
  \begin{center}
  $
  \begin{array}{l}
    \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
    \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
  \end{array}
  $
  \end{center}
  While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
  Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
  The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
  Thus, even if $\src{foo}$ may have been robust with respect to the \gls*{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
\end{example}

@@

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
% \MPpost{
%   this abstract fails in setting up the robustness argument.
%   it'll be a good exercise for you to take inspiration from this, but then write something as concise but with a more direct focus on robust properties
% }
% context
To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety as well as cryptographic constant time.
% need
Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
% task
Unfortunately, given two secure compiler passes that each preserve a possibly different security property, it is unclear what kind of security property is preserved by the composition of those secure compiler passes.
%there is no way to tell what kind of security property will the composition of those secure compilers preserve.

% object
This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
% findings
Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a theory of composition of secure compilers.
Then, it showcases this theory a secure multi-pass compiler that preserves the aforementioned security-relevant properties.
% conclusion
Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
% 
From an engineering perspective, this is the desirable approach to building secure compilers.

@@

\gls*{ms} is a security property obtained by composing \gls*{sms}, which ensures array accesses are all within bounds, and \gls*{tms}, which ensures pointers are only used when they are valid~\cite{azevedo2018meaningsofms,jim2002cyclone,necula2005ccured,nagarakatte2010cets,nagarakatte2009soft,akritidis2009baggy,michael2023mswasm}.
\gls*{cct} is a security property that ensures sensitive data is not leaked via timing side-channels~\cite{kocher1996timing}.
Together, \gls*{sms}, \gls*{tms} and \gls*{scct}, an enforceable overapproximation of \gls*{cct}, yield \gls*{msscct}, which is the gold standard of security properties for secure applications.
Programs attaining \gls*{msscct} do not leak sensitive data either through erroneous memory accesses, nor through timing side-channels.
As discussed in \Cref{ex:strncpy}, these security properties can be enforced by compiler passes~\cite{bond2017vale,almeida2017jasmin}, to ensure programmers need not be aware of the architectural details of where their code will run.

@@

Robust compilers~\cite{abate2019jour} are a form of secure compilers that preserve security properties even in the presence of arbitrary attackers interacting with compiled code.
Thus, robust compilers can be used to prevent vulnerabilities resulting from uninitialised memory (as well as many other ones), e.g., by targeting capability-based languages such as CHERI~\cite{woodruff2014CHERI}, Arm Morello~\cite{arm-morello}, or MSWasm~\cite{michael2023mswasm}, where the compiler relies on capabilities to check that pointers are always initialised.
\end{example}

@@

        % \MPpast{
        % 	some of the good things here: 
        % 	our definitions are amenable to sec comp proofs, unlike existing definitions
        % 	MS def is like mswasm MS def, 
        % 	but CCT is new
        % }

@@

Property satisfaction is defined on whole programs, i.e., programs without missing definitions.
Thus, from a security perspective, this considers only a passive attacker model, where the attacker observes the execution and, e.g., retrieves secrets from that.
To consider a stronger model similarly to what existing work does~\cite{abate2019jour,abate2021extacc,maffeis2008code-carrying,gordon2003authenticity,fournet2007authorization,bengtson2011refine,backes2014uniontyps,michael2023mswasm,swasey2017robust,sammler2019benefits}, the concept of satisfaction can be extended with {\em robustness}.
Robust satisfaction considers partial programs $\progvar$, i.e., components with missing imports, which cannot run until said imports are fulfilled.
To remedy this, {\em linking} takes two partial programs $\progvar[_{1}],\progvar[_{2}]$ and produces a whole program $\wholeprogvar$, i.e., $\link{\progvar[_{1}]}{\progvar[_{2}]}=\wholeprogvar$.
As typically done in works that consider the execution of partial programs~\cite{abate2019jour,devriese2018parametricity,patrignani2021rsc,korashy2021capableptrs,strydonck2019lincap,devriese2017modular,bowman2015noninterference,ahmed2011equivcps,patterson2017linkingtyps},
this paper assumes that whole programs are the result of linking partial programs referred to as {\em context} ($\ctx$) and {\em component} ($\comp$).
The context is an arbitrary program and thus has the role of an {\em attacker} that can interact with the component by means of whatever features the programming language has, and the component is what is security-relevant.
%In this work, the semantics of the programming language is expected to differentiate between the component and the context.
With this, \Thmref{def:propsat} can be extended as follows: for \bul{components $\progvar$ to robustly satisfy a property $\pi$}, take an \iul{attacker context $\contextvar$ and link it with $\progvar$}, \oul{the resulting whole program must satisfy $\pi$}.

@@

A {\em compiler} ($\cc{\src{L}}{\trg{L}}$) translates syntactic descriptions of programs from a {\em source} ($\src{L}$) into a {\em target} ($\trg{L}$) programming language.
This translation is considered {\em correct} if it is semantics-preserving.
That is, for a whole program $\src{\wholeprogvar}$, the compiler should relate the $\src{L}$ semantics of $\src{\wholeprogvar}$ with the semantics of $\trg{T}$ of the compiled counterpart of $\src{p}$ in such a way that they are ,,compatible''.
Unfortunately, correct compilers may be insecure compilers~\cite{patrignani2019survey,kennedy2006secure.net,abadi1999protect,ahmed2018dagstuhl} and programs translated by insecure compilers can violate security properties that the programmer assumes to hold.
To define when a compiler is secure, this paper uses the robust compilation framework~\cite{abate2019jour}, which the following definition summarises.

@@

\begin{example}[Types]
  Suppose $\src{L}$ is a statically-typed language similar to \texttt{C} and $\trg{T}$ is dynamically typed, where both share the same syntax up to dynamic type checks.
  Consider the following $\src{L}$ component and its compiled version below.
  \begin{center}
  $
  \begin{array}{l}
    \src{\lfunction{foo}{(char* x, int\ n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}} \\
    \trg{\lfunction{foo}{(\phantom{char*\ }x, \phantom{int\ }n)}{\lifz{valid\_ptr(x, n, sizeof(char))}{x[0]}{-1}}}
  \end{array}
  $
  \end{center}
  While the compiler emits code that may look correct, the generated code does not check that the provided argument is of the right type.
  Even though the pointer $\trg{x}$ is checked for validity, the context $\trg{foo((int*)y, 1)}$ is able to provoke a read out of bounds.
  Suppose the component transferred control to the context and passed ownership of a $\trg{char}$ pointer $\trg{y}$ sized $\trg{1}$ cells, the context can now call the component again, casting this buffer to an $\trg{int*}$ prior to that call.
  The pointer is valid for one $\trg{char}$-sized memory cell, as expected, but the actual read operation now returns $\trg{sizeof(int)}$ many bytes instead of just $\trg{sizeof(char)}$ many.
  Thus, even if $\src{foo}$ may have been robust with respect to the \gls*{sms}, its compiled counterpart is not and therefore the compiler fails to attain \Cref{def:rtp}.
\end{example}

@@

% \MPpost{
% 	i think the trace description of traces is more uindertandable,
% 	teh monitor one is easier to work with
% 	this is the justification for having both that we'll have to make in the future
% }
